---
title: "Analyzing German articles usage by children"
author: "Sadhana"
date: "2024-05-14"
output: html_document
---

# Creating main dataframe for the task 
### 1. Loading the necessary libraries. 


```{r}

library(childesr)
library(tidyverse)
library(ggplot2)
library(ggrepel)

```

### 2. Listing the functions in 'childesr' package

Since childes-db contains a set of tables, the 'childesr' package helps to retrieve and manipulate those tables with the help of below functions. For instance, get_collections() gives the names of available collections of corpora (“German”, “Spanish”, etc), and get_corpora() gives the names of available corpora.


```{r}

ls("package:childesr")

```


```{r}

collection <- get_collections()
head(collection)

```


```{r}

corpora <- get_corpora()
head(corpora)

```

### 3. Retrieving only the German data with the help of get_tokens() function

get_tokens() gives information about the words we need

```{r}

childes.Ger.article <- get_tokens(collection = "German", token = c("der","die","das","den","dem","ein","eine","einen","einem","einer"))

```
```{r}

head(childes.Ger.article)

```


### 4. Creating the main dataframe 
#### 4.1. Selecting only the required columns

```{r}

childes.Ger.article.df <- select(childes.Ger.article, c(gloss,speaker_role,target_child_name,target_child_age,corpus_name))
head(childes.Ger.article.df)

```

#### 4.2. Checking for NA values and removing those rows

```{r}

colSums(is.na(childes.Ger.article.df))

```

```{r}

childes.Ger.article.df <- na.omit(childes.Ger.article.df)

```


#### 4.3. Checking the list of German corpora 

```{r}

childes.Ger.article.df %>% count(corpus_name)

```

#### 4.4. Getting the range of ages of children (ages are represented in terms of months)

```{r}

min(childes.Ger.article.df$target_child_age)
max(childes.Ger.article.df$target_child_age)

```

#### 4.5. Selecting ages between 12 and 96 months (i.e., from 1 to 8 years)

```{r}

childes.Ger.article.df <- filter(childes.Ger.article.df, target_child_age >= 12 & target_child_age <= 96)
head(childes.Ger.article.df)

```

#### 4.6. Renaming the column name of 'gloss' to 'article'

```{r}

childes.Ger.article.df <- rename(childes.Ger.article.df, article=gloss)


```

```{r}

childes.Ger.article.df <- filter(childes.Ger.article.df, article %in% c("der","die","das","den","dem","ein","eine","einen","einem","einer"))

```

#### 4.7. Adding a new column called 'target_child_age_inyears' for representing ages in years


```{r}

childes.Ger.article.df <- mutate(childes.Ger.article.df, target_child_age = as.integer(target_child_age))

```


```{r}

childes.Ger.article.df <- mutate(childes.Ger.article.df, target_child_age_inyears = target_child_age / 12)

```

```{r}

childes.Ger.article.df <- mutate(childes.Ger.article.df, target_child_age_inyears = round(target_child_age_inyears, 1))
head(childes.Ger.article.df)

```

#### 4.8. Classifying ages in terms of stage (stage1, stage2, etc.) and adding it in a new column.

1 to 2 years belongs to stage1, 2 to 3 years belongs to stage2, 3 to 4 years belongs to stage3, 4 to 6 years belongs to stage4, and 6 to 8 years belongs to stage5

```{r}

childes.Ger.article.df %>% count(target_child_age_inyears)

```


```{r}

childes.Ger.article.df <-  mutate(childes.Ger.article.df, stage = case_when(target_child_age_inyears >=1.0 &  target_child_age_inyears <=2.0 ~ "stage1", target_child_age_inyears >=2.1 &  target_child_age_inyears <=3.0 ~ "stage2", target_child_age_inyears >=3.1 &  target_child_age_inyears <=4.0 ~ "stage3", target_child_age_inyears >=4.1 &  target_child_age_inyears <=6.0 ~ "stage4", target_child_age_inyears >=6.1 &  target_child_age_inyears <=8.0 ~ "stage5"))
            

```

```{r}

head(childes.Ger.article.df)


```
```{r}

summary(childes.Ger.article.df)

```

### 5. Saving the dataframe in excel format

```{r}

library(openxlsx)

```

```{r}

write.xlsx(childes.Ger.article.df, 'main_dataset.xlsx',colNames=TRUE, rowNames=TRUE)

```


# Research question1

At what age, children start to use articles in general? From when they start to use accusative and dative?

### 1. Creating a subset dataframe for research question1

#### 1.1. Copying the main dataframe

```{r}

childes.Ger.article.df1 <- childes.Ger.article.df

```


#### 1.2. Selecting only 'Target_Child' value in 'speaker_role' column, since this question is only related to children

```{r}

childes.Ger.article.df1 <- filter(childes.Ger.article.df1, speaker_role=="Target_Child")
head(childes.Ger.article.df1)

```


#### 1.3. Selecting only the required columns 

```{r}

childes.Ger.article.df1 <- select(childes.Ger.article.df1, c(article,stage))

head(childes.Ger.article.df1)

```

```{r}

summary(childes.Ger.article.df1)

```

#### 1.4. Saving the dataframe in excel format


```{r}

write.xlsx(childes.Ger.article.df1, 'Rq1_dataset.xlsx',colNames=TRUE, rowNames=TRUE)

```


### 2. MCA analysis for research question1


#### 2.1. Loading 'FactoMineR' package

```{r}

library(FactoMineR)

```


#### 2.3. Implementing MCA 

```{r}

mca.1 <- MCA(childes.Ger.article.df1, graph = FALSE)

mca.1

```


#### 2.4.  Summary of MCA

```{r}

summary(mca.1)

```



#### 2.5. Visualizing the graph of MCA

```{r}

plot(mca.1, invisible = "ind")

```

The plots of the above graph are very congested and overlapped. Hence, it is very difficult to do the interpretation. In order to overcome this problem, there is a package called 'factoextra' that provides an elegant and clear visualization of MCA graph without any overlapping. 


```{r}

library(factoextra)

```

'fviz_mca_var()' function is used to display the MCA graph of variable categories.

```{r}

fviz_mca_var(mca.1)

```


# Research question2 

Which articles' usage is high during childhood: Is it definite or indefinite?

### 1. Creating a subset dataframe for research question 2

#### 1.1. Copying the dataframe of research question 1

```{r}

childes.Ger.article.df2 <- childes.Ger.article.df1
head(childes.Ger.article.df2)

```

#### 1.2. Creating a new column for definite and indefinite articles

```{r}

childes.Ger.article.df2 <-  mutate(childes.Ger.article.df2, def = case_when(article %in% c("der","die","das","den","dem") ~ "definite", article %in% c("ein","eine","einen","einem","einer") ~ "indefinite" ))

```

#### 1.3. Selecting only required columns 

```{r}

childes.Ger.article.df2 <- select(childes.Ger.article.df2, c(stage,def))

head(childes.Ger.article.df2)

```

#### 1.4. Saving the dataframe

```{r}

write.xlsx(childes.Ger.article.df2, 'Rq2_dataset.xlsx',colNames=TRUE, rowNames=TRUE)

```


### 2. MCA analysis for research question 2

#### 2.1. Implementing MCA

```{r}

mca.2 <- MCA(childes.Ger.article.df2, graph = FALSE)

mca.2

```

#### 2.2. Summary of MCA

```{r}

summary(mca.2)

```


#### 2.3. Visualizing the graph of MCA


```{r}

plot(mca.2, invisible = "ind")

```



```{r}

ggplot(childes.Ger.article.df2, aes(x = def)) +
  geom_bar(stat = "count", aes(fill = def)) +
  labs(x = "definte / indefinite", y = "Frequency of definite & indefinite articles")

```

# Research question3

What is the overall usage of articles among children, is it more or less (by comparing it with adults' usage)?

### 1. Creating a dataframe for research question3

#### 1.1. Copying the main dataframe

```{r}

childes.Ger.article.df3 <- childes.Ger.article.df

```

#### 1.2. Getting the values of 'speaker_role' column

```{r}

unique(childes.Ger.article.df3$speaker_role)

```

#### 1.3. Removing the rows whose 'speaker_role' is not related to adults i.e., child, friend, playmate.

```{r}

childes.Ger.article.df3 <- filter(childes.Ger.article.df3, speaker_role != 'Child' & speaker_role != 'Friend' & speaker_role != 'Playmate')

```

#### 1.4. Adding new column called 'speaker'

```{r}

childes.Ger.article.df3 <-  mutate(childes.Ger.article.df3, speaker = case_when(speaker_role %in% c("Investigator","Mother"   ,"Grandfather", "Brother","Father","Relative","Grandmother","Visitor", "Adult", "Sister", "Teacher", "Sibling", "Unidentified", "Media" ) ~ "adults", speaker_role == "Target_Child" ~ "children"))

```

#### 1.5. Changing the values of 'article' column into definite and indefinite. 

Inorder to make the analysis clear and simple, the articles has been categorized into definite and indefinite instead of being der, die, das, etc.

```{r}

childes.Ger.article.df3 <-  mutate(childes.Ger.article.df3, article = case_when(article %in% c("der","die","das","den","dem") ~ "definite", article %in% c("ein","eine","einen","einem","einer") ~ "indefinite" ))

```


#### 1.6. Selecting only the required columns

```{r}

childes.Ger.article.df3 <- select(childes.Ger.article.df3, c(article,speaker))
head(childes.Ger.article.df3)

```

#### 1.7. Saving the dataframe

```{r}

write.xlsx(childes.Ger.article.df3, 'Rq3_dataset.xlsx',colNames=TRUE, rowNames=TRUE)

```


### 2. MCA analysis for research question3

#### 2.1. Implementing MCA

```{r}

mca.3 <- MCA(childes.Ger.article.df3, graph = FALSE)
mca.3

```

#### 2.2 Summary of MCA

```{r}

summary(mca.3)

```


#### 2.3. Visualizing the graph of MCA

```{r}

plot(mca.3, invisible = "ind")

```



```{r}

ggplot(childes.Ger.article.df3, aes(x = speaker)) +
       geom_bar(stat = "count") +
   facet_wrap(~article) +
  labs(x = "adults vs children", y = "overall usage of articles") +
    theme(axis.text.x = element_text(angle = 90))


```

#### checking the count of definite and indefinite articles against speaker(adults and children)


```{r}

count <- childes.Ger.article.df3 %>% count(speaker, article)
count

```


## Additional commands

### Retrieving transcripts of 'den' for stage3


```{r}

den.utterence_stage3 <- get_utterances(collection = "German", age = c(36, 48))

```

```{r}

den.utterence_stage3 <- select(den.utterence_stage3, c(id, gloss,speaker_role,target_child_name,target_child_age,corpus_name))

```

```{r}

den.utterence_stage3 <- filter(den.utterence_stage3, grepl("\\bden\\b", gloss, ignore.case = TRUE))

```


```{r}

den.utterence_stage3 <- filter(den.utterence_stage3, corpus_name == "Miller" )

write.xlsx(den.utterence_stage3, 'den_uttr_stage3.xlsx',colNames=TRUE, rowNames=TRUE)

```


### Retrieving transcripts of 'dem' for stage3


```{r}

dem.utterence_stage3 <- get_utterances(collection = "German", age = c(36, 48))

```

```{r}

dem.utterence_stage3 <- select(dem.utterence_stage3, c(id, gloss,speaker_role,target_child_name,target_child_age,corpus_name))

```

```{r}

dem.utterence_stage3 <- filter(dem.utterence_stage3, grepl("\\bdem\\b", gloss, ignore.case = TRUE))

```

```{r}

dem.utterence_stage3 <- filter(dem.utterence_stage3, corpus_name == "Miller" )

write.xlsx(dem.utterence_stage3, 'dem_uttr_stage3.xlsx',colNames=TRUE, rowNames=TRUE)

```




